Tango-cc analysis

### 2024-01-31 stats/analysis.Rmd
script for server

# color palette
# qual_col_pals = brewer.pal.info[brewer.pal.info$colorblind == TRUE,]
# col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
# 
# # kelly palette of 22 colors maximize the contrast between colors in a set if the colors are chosen in sequential order. Kelly paid attention to the needs of people with color blindness
# colors=kelly(n=n_distinct(data$community)+1)

@@

```{r load_data}
data <- read_csv("../data/data.csv") %>% 
    mutate(continent = ifelse(
    community %in% c("leipzig", "plymouth"), "Europe", 
    ifelse(
      community %in% c("akure", "hai||om", "khwe", "chimfunshi", "bandongo", "bayaka", "windhoek", "uganda"), "Africa",
      ifelse(community %in% c("stanford", "mexico", "buenos_aires"), "America",
             ifelse(community %in% c("india", "beijing", "turkey"), "Asia", "Oceania"))
    )
  ))%>%
  mutate(continent = factor(continent, levels = c("America", "Africa", "Europe", "Asia", "Oceania"), ordered = T))%>%
  mutate(community = recode(community,
                            akure = "Akure (Nigeria)",
                            leipzig = "Leipzig (Germany)",
                            `hai||om` = "Hai||om (Namibia)",
                            khwe = "Khwe (Namibia)",
                            windhoek = "Windhoek (Namibia)",
                            stanford = "Stanford (USA)",
                            chimfunshi = "Chimfunshi (Zambia)",
                            mexico = "Ocuilan (Mexico)",
                            plymouth = "Plymouth (UK)",
                            beijing = "Beijing (China)", 
                            india = "Pune (India)",
                            buenos_aires = "Buenos Aires (Argentina)",
                            auckland = "Auckland (New Zealand)",
                            turkey = "Malatya (Türkiye)",
                            bandongo = "Bandongo (Rep. Congo)",
                            bayaka = "BaYaka (Rep. Congo)", 
                            uganda = "Nyabyeya (Uganda)")
  ) %>% 
  mutate(
    click = abs(clickdistfromtargetcenterx)/160,
    mean_click = mean(abs(clickdistfromtargetcenterx) / 160),
    .by = c(ageinyears, community, subjid)
  )

mdata <- data %>%
  filter(trialtype == "test",
         voiceover == F
         )%>%
  mutate(trialnr = as.numeric(factor(trialnr)))%>%
  group_by(subjid, community, age_group, ageinyears)%>%
  summarise(impr = mean(abs(clickdistfromtargetcenterx)))%>%
  ungroup()%>%
  mutate(age = scale(ageinyears))
```

```{r sample}
# full sample
data %>%
  group_by(subjid) %>%
  summarise(n= n())%>%
  arrange(n)

data %>%
  group_by(community, age_group) %>%
  summarise(n= n_distinct(subjid))

data %>%
  group_by(community) %>%
  summarise(n= n_distinct(subjid))

data %>%
  summarise(n= n_distinct(subjid))

@@

# Internal consistency

```{r exclude_children}
# children who did not finish all trials are removed (only 5 children total)
exclude <- data %>%
  filter(trialtype == "test", voiceover == F) %>%
  group_by(subjid) %>%
  summarise(n= n())%>%
  filter(n < 15)%>%
  pull(subjid)

reli_data <- data%>%
  filter(trialtype == "test", voiceover == F) %>%
  filter(!subjid %in% exclude)%>%
  mutate(pos_cent = as.character(abs(5-targetposition)))
```

## Odd-even

```{r odd_even_plot}
reli_data%>%
  mutate(half = ifelse(trialnr %% 2 == 0, "half_1", "half_2"))%>%
  group_by(subjid, community, half)%>%
  summarise(impr = mean(abs(clickdistfromtargetcenterx)))%>%
  pivot_wider(names_from = half, values_from = impr)%>%
  group_by(community)%>%
  summarise(reli = cor(half_1, half_2))%>%
  ggplot(aes(x = reli, y = community))+
  geom_vline(xintercept = 0, lty = 3, alpha = .75)+
  geom_point(alpha = .5)+
  # geom_pointrange(aes(xmin = lci, xmax = uci))+
  scale_x_continuous(limits = c(-0.5, 1), breaks = c(-0.5, 0, 0.5, 0.6,0.7, 0.8, 0.9, 1))+
  ggtitle("Simple odd-even split")+
  theme_minimal()
```

@@

```{r stratified_targetposition_data}
# fn_mean <- function(ds) {
#   return(mean(abs(ds$clickdistfromtargetcenterx), na.rm = TRUE))
# }
# 
# #rel <- tibble()
# 
# communities <- unique(data$community)
# 
# for (j in c(21:50)) {
# 
# for (i in communities) {
#   
# # stratified by target position
# rel_s <- by_split(
#   data = reli_data %>% filter(community == i),
#   participants = reli_data %>% filter(community == i) %>% pull(subjid),
#   fn_score = fn_mean,
#   verbose = F,
#   stratification = paste(reli_data %>% filter(community == i) %>% pull(pos_cent)),
#   careful = FALSE
# )
# 
# samp <- tibble(
#   community = i,
#   iter = j,
#   reli = split_coefs(rel_s, cor),
#   lci = split_ci(rel_s, cor)$lims["0.025","bca"],
#   uci = split_ci(rel_s, cor)$lims["0.975","bca"]
#   )
# 
# rel <- bind_rows(rel, samp)
# 
# }
# 
# }
# 
# saveRDS(rel, "../saves/stratified_internal_consistency.rds")

rel <- readRDS("../saves/stratified_internal_consistency.rds")
```

```{r stratified_targetposition_plot}
ggplot(rel, aes(x = reli, y = community))+
  geom_vline(xintercept = 0, lty = 3, alpha = .75)+
  geom_density_ridges(col = "white", alpha = .5,rel_min_height = 0.01,)+
  geom_point(alpha = .5)+
  #geom_pointrange(aes(xmin = lci, xmax = uci))+
  scale_x_continuous(limits = c(-0.5, 1), breaks = c(-0.5, 0, 0.5, 0.6,0.7, 0.8, 0.9, 1))+
  ggtitle("Stratified by target position - no age correction")+
  theme_minimal()
```

@@

```{r stratified_agecorrected_data}
# # estimate age-corrected reliability
# # run model first for one community to use update()
# split <- split_df(
#   data = reli_data %>% filter(community == "leipzig"),
#   stratification = reli_data %>% filter(community == "leipzig") %>% pull(pos_cent),
#   careful = FALSE
# )
# 
# sdata <- bind_rows(
#   split[[1]]%>%select(subjid, ageinyears, clickdistfromtargetcenterx)%>%mutate(impr = abs(clickdistfromtargetcenterx), half = "half_1"),
#   split[[2]]%>%select(subjid, ageinyears, clickdistfromtargetcenterx)%>%mutate(impr = abs(clickdistfromtargetcenterx), half = "half_2"))%>%
#   mutate(age = scale(ageinyears))
# 
# reli_age <- brm(impr ~ age + (0 + half | subjid), 
#                 data   = sdata, 
#                 family = lognormal(), 
#                 warmup = 1000, 
#                 iter   = 3000,
#                 control = list(adapt_delta = 0.95),
#                 chains = 4,
#                 cores  = 4)
# 
# # loop over communities
# 
# communities <- unique(data$community)
# 
# #model_cor <- tibble()
# 
# for (j in c(21:50)) {
#   
# for (i in communities) {
#   
# com_split <- split_df(
#   data = reli_data %>% filter(community == i),
#   stratification = reli_data %>% filter(community == i) %>% pull(pos_cent),
#   careful = FALSE
#   )
# 
# com_sdata <- bind_rows(
#   com_split[[1]]%>%select(subjid, ageinyears, clickdistfromtargetcenterx)%>%mutate(impr = abs(clickdistfromtargetcenterx), half = "half_1"),
#   com_split[[2]]%>%select(subjid, ageinyears, clickdistfromtargetcenterx)%>%mutate(impr = abs(clickdistfromtargetcenterx), half = "half_2"))%>%
#   mutate(age = scale(ageinyears))
#  
# com_mod <-  update(reli_age,
#         newdata = com_sdata,
#         warmup = 1000, 
#         iter   = 3000,
#         control = list(adapt_delta = 0.95),
#         chains = 4,
#         cores  = 4)
#   
#   cor <- summary(com_mod)$random$subjid%>%as_tibble(rownames = "parameter")%>%filter(parameter == "cor(halfhalf_1,halfhalf_2)")%>%mutate(community = i, iter = j)
# 
#   model_cor <- bind_rows(model_cor, cor)
#   
#   
#   }
#   
# }
# 
# saveRDS(model_cor, "../saves/age_corrected_internal_consistency.rds")

model_cor <- readRDS("../saves/age_corrected_internal_consistency.rds")  
```

```{r stratified_agecorrected_plot}
ggplot(model_cor, aes(x = Estimate, y = community))+
  geom_vline(xintercept = 0, lty = 3, alpha = .75)+
  geom_density_ridges(col = "white", alpha = .5, rel_min_height = 0.01,)+
  geom_point(alpha = .5)+
  #geom_pointrange(aes(xmin = `l-95% CI`, xmax = `u-95% CI`), alpha = .5)+
  ggtitle("Stratified by target position - model-based age correction")+
  scale_x_continuous(limits = c(-1, 1), breaks = c(-0.5, 0, 0.5, 0.6,0.7, 0.8, 0.9, 1))+
  theme_minimal()
```

```{r beautified_plot}
model_cor <- model_cor %>% 
    mutate(continent = ifelse(
    community %in% c("leipzig", "plymouth"), "Europe", 
    ifelse(
      community %in% c("akure", "hai||om", "khwe", "chimfunshi", "bandongo", "bayaka", "windhoek", "uganda"), "Africa",
      ifelse(community %in% c("stanford", "mexico", "buenos_aires"), "America",
             ifelse(community %in% c("india", "beijing", "turkey"), "Asia", "Oceania"))
    )
  ))%>%
  mutate(continent = factor(continent, levels = c("America", "Africa", "Europe", "Asia", "Oceania"), ordered = T))%>%
  mutate(community = recode(community,
                            akure = "Akure (Nigeria)",
                            leipzig = "Leipzig (Germany)",
                            `hai||om` = "Hai||om (Namibia)",
                            khwe = "Khwe (Namibia)",
                            windhoek = "Windhoek (Namibia)",
                            stanford = "Stanford (USA)",
                            chimfunshi = "Chimfunshi (Zambia)",
                            mexico = "Ocuilan (Mexico)",
                            plymouth = "Plymouth (UK)",
                            beijing = "Beijing (China)", 
                            india = "Pune (India)",
                            buenos_aires = "Buenos Aires (Argentina)",
                            auckland = "Auckland (New Zealand)",
                            turkey = "Malatya (Türkiye)",
                            bandongo = "Bandongo (Rep. Congo)",
                            bayaka = "BaYaka (Rep. Congo)", 
                            uganda = "Nyabyeya (Uganda)")
  )

model_cor %>% 
  ggplot(., aes(x = Estimate, y = community))+
  geom_vline(xintercept = 0, lty = 3, alpha = .75)+
  geom_density_ridges(col = "white", alpha = .5, rel_min_height = 0.01, fill="#29485D")+
  geom_point(alpha = .5)+
  #geom_pointrange(aes(xmin = `l-95% CI`, xmax = `u-95% CI`), alpha = .5)+
  # ggtitle("Stratified by target position - model-based age correction")+
  scale_x_continuous(limits = c(-1, 1), breaks = c(-1, -0.5, 0, 0.5, 0.6,0.7, 0.8, 0.9, 1))+
  theme_light() +
  labs(y = "", x = "Splithalf reliability estimate", caption = "(stratified by target position; model-based age correction)")

@@

```{r scatterplot}
reli_data %>% 
  ggplot() +
  geom_point(aes(x = ageinyears, y = mean_click, col = community), alpha = 0.9) +
  scale_color_manual(values = custom_palette) +
  scale_x_continuous(limits = c(2, 11), breaks = seq(2, 11, 1)) +
  scale_y_continuous(limits = c(0, 7), breaks = seq(0, 7, 1)) +
  labs(x = "Age in years", y = "Average absolute imprecision in target width", col = "")

>>>>

# color palette
# qual_col_pals = brewer.pal.info[brewer.pal.info$colorblind == TRUE,]
# col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
# 
# # kelly palette of 22 colors maximize the contrast between colors in a set if the colors are chosen in sequential order. Kelly paid attention to the needs of people with color blindness
# colors=kelly(n=n_distinct(data$community) +1)

@@

```{r load_data}
complete_data <- read_csv("../data/data.csv")
```

```{r data_preparation}
# complete dataset, with all trials
all_trials <- complete_data %>% 
    mutate(continent = ifelse(
    community %in% c("leipzig", "plymouth"), "Europe", 
    ifelse(
      community %in% c("akure", "hai||om", "khwe", "chimfunshi", "bandongo", "bayaka", "windhoek", "uganda"), "Africa",
      ifelse(community %in% c("stanford", "mexico", "buenos_aires"), "Americas",
             ifelse(community %in% c("india", "beijing", "turkey"), "Asia", "Oceania"))
    )
  )) %>%
  mutate(continent = factor(continent, levels = c("Americas", "Africa", "Europe", "Asia", "Oceania"), ordered = T)) %>%
  mutate(community = recode(community,
                            akure = "Akure (Nigeria)",
                            leipzig = "Leipzig (Germany)",
                            `hai||om` = "Hai||om (Namibia)",
                            khwe = "Khwe (Namibia)",
                            windhoek = "Windhoek (Namibia)",
                            stanford = "Stanford (USA)",
                            chimfunshi = "Chimfunshi (Zambia)",
                            mexico = "Ocuilan (Mexico)",
                            plymouth = "Plymouth (UK)",
                            beijing = "Beijing (China)", 
                            india = "Pune (India)",
                            buenos_aires = "Buenos Aires (Argentina)",
                            auckland = "Auckland (New Zealand)",
                            turkey = "Malatya (Türkiye)",
                            bandongo = "Bandongo (Rep. Congo)",
                            bayaka = "BaYaka (Rep. Congo)", 
                            uganda = "Nyabyeya (Uganda)")
  ) %>% 
  mutate(
    imprecision = abs(clickdistfromtargetcenterx)/160,
    centrality = as.character(abs(5-targetposition))
  )

# dataset with test trials only 
data <- all_trials %>%
  filter(trialtype == "test", voiceover == F) %>%
  mutate(trialnr = as.numeric(factor(trialnr)))

# dataset with test trials only, with mean imprecision per participant
mdata <- data %>%
  summarise(mean_imprecision = mean(imprecision), .by = c(subjid, community, age_group, ageinyears)) %>%
  mutate(age = scale(ageinyears))
```

```{r sample}
# full sample
all_trials %>%
  group_by(subjid) %>%
  summarise(n= n()) %>%
  arrange(n)

all_trials %>%
  group_by(community, age_group) %>%
  summarise(n= n_distinct(subjid))

all_trials %>%
  group_by(community) %>%
  summarise(n= n_distinct(subjid))

all_trials %>%
  summarise(n= n_distinct(subjid))

@@

```{r exclude_children}
# children who did not finish all trials are removed (only 5 children total)
exclude <- all_trials %>%
  filter(trialtype == "test", voiceover == F) %>%
  group_by(subjid) %>%
  summarise(n= n()) %>%
  filter(n < 15) %>%
  pull(subjid)

all_trials <- all_trials %>% filter(!subjid %in% exclude)
data <- data %>% filter(!subjid %in% exclude)
mdata <- mdata %>% filter(!subjid %in% exclude)
```

# Internal Consistency
## Odd-even

```{r odd_even_plot}
data %>%
  mutate(half = ifelse(trialnr %% 2 == 0, "half_1", "half_2")) %>%
  group_by(subjid, community, half) %>%
  summarise(mean_imprecision = mean(imprecision)) %>%
  pivot_wider(names_from = half, values_from = mean_imprecision) %>%
  group_by(community) %>%
  summarise(reli = cor(half_1, half_2)) %>%
  ggplot(aes(x = reli, y = community)) +
  geom_vline(xintercept = 0, lty = 3, alpha = .75) +
  geom_point(alpha = .5) +
  scale_x_continuous(limits = c(-0.5, 1), breaks = c(-0.5, 0, 0.5, 0.6,0.7, 0.8, 0.9, 1)) +
  labs(x = "Simple odd-even split") +
  theme_minimal()
```

@@

```{r stratified_targetposition_data}
fn_mean <- function(ds) {
  return(mean(abs(ds$clickdistfromtargetcenterx), na.rm = TRUE))
}

rel <- tibble()

communities <- unique(data$community)

for (j in c(1:50)) {
  for (i in communities) {
    # stratified by target position
    rel_s <- by_split(
      data = data %>% filter(community == i),
      participants = data %>% filter(community == i) %>% pull(subjid),
      fn_score = fn_mean,
      verbose = F,
      stratification = paste(data %>% filter(community == i) %>% pull(centrality)),
      careful = FALSE
      )

    samp <- tibble(
      community = i,
      iter = j,
      reli = split_coefs(rel_s, cor),
      lci = split_ci(rel_s, cor)$lims["0.025","bca"],
      uci = split_ci(rel_s, cor)$lims["0.975","bca"]
      )

    print(i)
    print(j)
    print(rel_s)
    print(samp)

    rel <- bind_rows(rel, samp)
  }
}

saveRDS(rel, "../saves/stratified_internal_consistency.rds")

# rel <- readRDS("../saves/stratified_internal_consistency.rds")
```

```{r stratified_targetposition_plot}
ggplot(rel, aes(x = reli, y = community)) +
  geom_vline(xintercept = 0, lty = 3, alpha = .75) +
  geom_density_ridges(col = "white", alpha = .5,rel_min_height = 0.01,) +
  geom_point(alpha = .5) +
  #geom_pointrange(aes(xmin = lci, xmax = uci)) +
  scale_x_continuous(limits = c(-0.5, 1), breaks = c(-0.5, 0, 0.5, 0.6,0.7, 0.8, 0.9, 1)) +
  ggtitle("Stratified by target position - no age correction") +
  theme_minimal()
```

@@

```{r stratified_agecorrected_data}
# estimate age-corrected reliability
# run model first for one community to use update()
split <- split_df(
  data = data %>% filter(community == "Leipzig (Germany)"),
  stratification = data %>% filter(community == "Leipzig (Germany)") %>% pull(centrality),
  careful = FALSE
)

sdata <- bind_rows(
  split[[1]] %>% 
    select(subjid, ageinyears, clickdistfromtargetcenterx) %>% 
    mutate(impr = abs(clickdistfromtargetcenterx), half = "half_1"),
  split[[2]] %>% 
    select(subjid, ageinyears, clickdistfromtargetcenterx) %>%
    mutate(impr = abs(clickdistfromtargetcenterx), half = "half_2")) %>%
  mutate(age = scale(ageinyears))

reli_age <- brm(impr ~ age + (0 + half | subjid),
                data   = sdata,
                family = lognormal(),
                warmup = 1000,
                iter   = 3000,
                control = list(adapt_delta = 0.95),
                chains = 4,
                cores  = 4)

# loop over communities
communities <- unique(data$community)

model_cor <- tibble()

for (j in c(1:50)) {
  for (i in communities) {

    com_split <- split_df(
      data = data %>% filter(community == i),
      stratification = data %>% filter(community == i) %>% pull(centrality),
      careful = FALSE
    )

    com_sdata <- bind_rows(
      com_split[[1]] %>% 
        select(subjid, ageinyears, clickdistfromtargetcenterx) %>%
        mutate(impr = abs(clickdistfromtargetcenterx), half = "half_1"),
      com_split[[2]] %>%
        select(subjid, ageinyears, clickdistfromtargetcenterx) %>%
        mutate(impr = abs(clickdistfromtargetcenterx), half = "half_2")) %>%
      mutate(age = scale(ageinyears))

    com_mod <- update(reli_age,
        newdata = com_sdata,
        warmup = 1000,
        iter   = 3000,
        control = list(adapt_delta = 0.95),
        chains = 4,
        cores  = 4)

    cor <- summary(com_mod)$random$subjid %>%
      as_tibble(rownames = "parameter") %>%
      filter(parameter == "cor(halfhalf_1,halfhalf_2)") %>%
      mutate(community = i, iter = j)

    model_cor <- bind_rows(model_cor, cor)
  }
}

saveRDS(model_cor, "../saves/age_corrected_internal_consistency.rds")

# model_cor <- readRDS("../saves/age_corrected_internal_consistency.rds")  
```

```{r stratified_agecorrected_plot}
ggplot(model_cor, aes(x = Estimate, y = community)) +
  geom_vline(xintercept = 0, lty = 3, alpha = .75) +
  geom_density_ridges(col = "white", alpha = .5, rel_min_height = 0.01,) +
  geom_point(alpha = .5) +
  #geom_pointrange(aes(xmin = `l-95% CI`, xmax = `u-95% CI`), alpha = .5) +
  ggtitle("Stratified by target position - model-based age correction") +
  scale_x_continuous(limits = c(-1, 1), breaks = c(-0.5, 0, 0.5, 0.6,0.7, 0.8, 0.9, 1)) +
  theme_minimal()
```

### run until here

```{r beautified_plot}
model_cor <- model_cor %>% 
    mutate(continent = ifelse(
    community %in% c("leipzig", "plymouth"), "Europe", 
    ifelse(
      community %in% c("akure", "hai||om", "khwe", "chimfunshi", "bandongo", "bayaka", "windhoek", "uganda"), "Africa",
      ifelse(community %in% c("stanford", "mexico", "buenos_aires"), "Americas",
             ifelse(community %in% c("india", "beijing", "turkey"), "Asia", "Oceania"))
    )
  )) %>%
  mutate(continent = factor(continent, levels = c("Americas", "Africa", "Europe", "Asia", "Oceania"), ordered = T)) %>%
  mutate(community = recode(community,
                            akure = "Akure (Nigeria)",
                            leipzig = "Leipzig (Germany)",
                            `hai||om` = "Hai||om (Namibia)",
                            khwe = "Khwe (Namibia)",
                            windhoek = "Windhoek (Namibia)",
                            stanford = "Stanford (USA)",
                            chimfunshi = "Chimfunshi (Zambia)",
                            mexico = "Ocuilan (Mexico)",
                            plymouth = "Plymouth (UK)",
                            beijing = "Beijing (China)", 
                            india = "Pune (India)",
                            buenos_aires = "Buenos Aires (Argentina)",
                            auckland = "Auckland (New Zealand)",
                            turkey = "Malatya (Türkiye)",
                            bandongo = "Bandongo (Rep. Congo)",
                            bayaka = "BaYaka (Rep. Congo)", 
                            uganda = "Nyabyeya (Uganda)")
  )

model_cor %>% 
  ggplot(., aes(x = Estimate, y = community)) +
  geom_vline(xintercept = 0, lty = 3, alpha = .75) +
  geom_density_ridges(col = "white", alpha = .5, rel_min_height = 0.01, fill="#29485D") +
  geom_point(alpha = .5) +
  #geom_pointrange(aes(xmin = `l-95% CI`, xmax = `u-95% CI`), alpha = .5) +
  # ggtitle("Stratified by target position - model-based age correction") +
  scale_x_continuous(limits = c(-1, 1), breaks = c(-1, -0.5, 0, 0.5, 0.6,0.7, 0.8, 0.9, 1)) +
  theme_light() +
  labs(y = "", x = "Splithalf reliability estimate", caption = "(stratified by target position; model-based age correction)")

@@

```{r scatterplot}
data %>% 
  ggplot() +
  geom_point(aes(x = ageinyears, y = mean_click, col = community), alpha = 0.9) +
  scale_color_manual(values = custom_palette) +
  scale_x_continuous(limits = c(2, 11), breaks = seq(2, 11, 1)) +
  scale_y_continuous(limits = c(0, 7), breaks = seq(0, 7, 1)) +
  labs(x = "Age in years", y = "Average absolute imprecision in target width", col = "")
